---
title: "Felicitas-Ruiz_Capstone_R"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#TITLE: Are there gender differences between and within each carcinoma disease group?

#About the data:
The question that I wanted to ask with R was an age question between males and females. I wanted to add age column to my dataset for problem 1 in R. But I didn't know quite how to do it in R so I did it in python. I then saved the new datset that now has a new column to a ['Filtered_edit_df.csv'](data/Filtered_edit_df.csv). I used this file to answer my question below

#Questions that I will ask with R
1. The authors noted that females with adenocarcinoma of the pancreato-biliary system had a lower median overall survival (months) than males (Table 2 in their paper). One question I wanted to ask was if females with Hepatocellular Cancer, so the control cancer, had a lower overall survival than males.

#Assumptions
Since this dataset is an observational study and there may be patients that are still alive, I do make a major assumption. If the date_death column is does not have a date, then I assume that person is still alive. Since this data was last assessed in August, 2016I assume the date is 01. I use these assumptions to calculate the age of these individuals in python and add that age to a new column. This dataset is then processed in R to answer the question above.
##Load tidyverse first
```{r}

library(tidyverse)
```
##Then I will configure file names
```{r}
Filtered <- read.csv(file = "data/Filtered_edit_df.csv")

Filtered
"Filtered datset currently 391 Rows"
```
#1. The authors noted that females with adenocarcinoma of the pancreato-biliary system had a lower median overall survival (months) than males (Table 2 in their paper). One question I wanted to ask was if females with Hepatocellular Cancer, so the control cancer, had a lower overall survival than males.
```{r}
# You are reading in the csv file into a tibble assigned as 'Filtered'.
Filtered2 <- Filtered %>%
  #Using pipe %>% operator to chain commands. Select() command to keep only the columns: pad (to later sort by pathological diagnosis), gender and chemo (becuase I am interesed in the discrepancy between gender and chemo).
  select(pad_only, gender, Age..in.days.) 
Filtered3 <- Filtered2[grep("Hepatocellular", Filtered2$pad_only), ] 

#Here I wanted to arrange by age with the eldest age (in days) at the top
Filtered5 <-Filtered3 %>%
  arrange(-(Age..in.days.)) %>%
  #I wanted to change the name of the columns
  mutate(Age_in_days=Age..in.days., Tumor_sample=pad_only) %>%
  select(Tumor_sample, gender, Age_in_days) %>%
 
  print()
```
```{r}
#Here I just wanted to subset the 10 eldest people in the Hepatocellular Carcinoma group
Filtered4 <-Filtered3 %>%
  arrange(-(Age..in.days.)) %>%
  slice(0:10) %>%
  mutate(Age_in_days=Age..in.days., Tumor_sample=pad_only) %>%
  select(Tumor_sample, gender, Age_in_days) %>%
  #count(gender)%>%
  #Now you are printing what you have selected and filtered with the chained commands. 
  print()


#gender
#f	13			
#m	62

```
#I find just by visually looking at the table that is printed from 'Filter4' that there are more males in the eldest group of 10 people that were found to have Hepatocellular Cancer. This was done by looking at the "Age_in_days" column I see that there are more males than females in the older age group. One way that we can visuLize this is by using ggplot down below. We see that in the eldest 10 patients ggplot, we see that males dominate are the predominate gender in this subset. In entire Hepatocellular Cancer group, it is less appareant that males live longer than females beucase there are more males enrolled in this Hepatocellular Cancer group cohort study than females. This is could be a confounding variable in the authors study as this control group consists of males. It would be interesting to learn more about the demographics of this cohort study to drawing more analysis.

```{r}
#Now I will make a plot to show the age distribution (in days) of patients in the Hepatocellular Cancer group based on gender but only the 10 eldest folks
plot_with_subset <- ggplot(Filtered4, aes(x = gender,
                 y = Age_in_days)) +
  geom_point()+
  theme_classic()+
  labs(x = "Gender of patient", y= "Age of patient (in days)", title="Age Distribution of Hepatocellular Cancer based on gender (eldest 10 patients)")

plot_with_subset

#Now I will make a plot to show the age distribution (in days) of patients in the Hepatocellular Cancer group based on gender (all data) 
plot_with_all <- ggplot(Filtered5, aes(x = gender,
                 y = Age_in_days)) +
  theme_classic()+
  geom_point()+
  labs(x = "Gender of patient", y= "Age of patient (in days)", title="Age Distribution of Hepatocellular Cancer based on gender (all data)")
plot_with_all
```

#Visualization Justification
I chose to use ggplot() with geom_point() to iluustrate the distribution of ages. I beieve that a histogram would limit the analysis but geom_point() would clearly depict all the points. 

#Reproducibility
Since I just both R and python, I do believe my analysis is reproducibile. I'm sure there are ways to calculate age in R but I wasn't sure. There might be a more conveient way than using both R and python to answer this question.



