---
title: "Feli Capstone Project README.md"
output: html_document

---
#output: 
#  md_document:
#     variant: markdown_github
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
library(tidyverse)
library(readr)
```


## Project organization

Information about things in this project

Within the data folder, I have placed the dataset that I extracted from DRYAD. The dataset had a "pcbil_clinicaldata.xlsx" excelsheet but in order to make dtaaframes in Python, I made separate CSV files for each sheet in the excel workbook. I also made my own csv files that I modified from "pcbil_clinicaldata.xlsx" csv files. These files are found within the [data folder](data/).

Scripts.All of my code, whether that be R.Rmd files or the python.ipynb files are found readily in my repository.


##About the Data

The dataset that I chose to investigate further was the [Immunohistochemical typing of adenocarcinomas](https://datadryad.org/stash/dataset/doi:10.5061/dryad.g8h71).
I am specifically working with the dataset that was found in the raw folder in "pcbil_clinicaldata.xlsx". This can be found in my repository folder at  ["pcbil_clinicaldata.xlsx"](data/pcbil_clinicaldata.xlsx). But since there are multiple tabs and I need to work with csv files, I will specifically be working with data that is found at ["pcbil_clinicaldata.xlsx"](data/pcbil_clinicaldata.xlsx) but in the "PLOS Revision" tab. I have placed the data that was in the "PLOS Revision" tab into a new csv file found [here](data/pcbil_clinicaldata_PLOS_REVISION.csv). 


The paper delves into Adenocarcinomas of the pancreatobiliary system and how they are classified. This data comes from an observational cohort study on patients with Adenocarcinoma of the pancreatobiliary system. Certain morphological features of carcinomas make it difficult to distinguish between carcinomas. For instance intrahepatic cholangiocarcinoma is difficult to distinguish based on morphology so other factors such as immunhistochemistry may be used. 


One assumptions I made about the data was the last time it was assessed. This is an observationaly data set that followed patients until August 2016. One of my questions involved age and so I assumed if the date od death was not listed, then they were still alive at least until the last time the data was assessed. I'm not sure exactly what day in August 2016 they stopped being followed but I listed the "date of death" as the last day the data was assessed just so that I can calculate Age based on days. So I did change the cells from "NaN" to present so that I could calculate the age. This edited csv file can be found [here](data/pcbil_clinicaldata_PLOS_REVISION_edited.csv).

Since I am using the datset that has already been precoessed by the group, I will describe a little what pre-processing the group did. They had originally used 38 antibody markers but reduced it down to 27 due to missing data. These markers are used commonly for diagnosing pathology such as markers for mucins to be able to distinguish between squamous and glandular and so.
Additionally, they imputed the dataset with the Bioconductor package “impute”. My analysis will be on the imputed dataset found [here](data/pcbil_clinicaldata_PLOS_REVISION_edited.csv) as previously stated.



## My three research questions to answer:

1. The authors noted that females with adenocarcinoma of the pancreato-biliary system had a lower median overall survival (months) than males (Table 2 in their paper). One question I wanted to ask was if females with Hepatocellular Cancer, so the control cancer, had a lower overall survival than males.

2. Continung with the theme of gender differences within disease groups, I wanted to see if there are also differences in care. For instance, are males more likely to receive chemotherapy than females in each adenocarcinoma disease group and in the hepatocellular carcinoma group? I will answer this question in Python.

3. Continung with the theme of gender differences in this dataset, I wanted to investigate if there are gender preferences when using a specific diagnostic probe. Is the diagnostic probe biopsy more often used on females than males? if there are preferences, do these preferences hold with different disease groups. I will answer this question in Python.


##Reproducibility
I believe I have given good comments to my code along the way so that my analysis is reproducible. Though the results may change given different patient samples. The dataset that was used often had more samples from males than females. For instance, in the hepatocellular carcinoma group there are more males than females and this may affect the analysis of results. A higher and equal distribution of males and females may affect the results. Also, additional information about patient demographics may also affect the reproducibility of the data.


